---
title: "Appendix 1"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Introduction

In this appendix, we carry out a small simulation study to assess the bias in the parameter estimates of a capture-recapture model to infer social networks.

## 2. Code writing for simulation and estimation

First, we load the `R2jags` package that will be used to fit models.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
library(R2jags)
```

Then we write a function `sim_CRnetwork` to simulate data. Note that it is used in `Jags`, as explained [here](https://oliviergimenez.github.io/post/sim_with_jags/).

```{r message=FALSE, warning=FALSE, paged.print=FALSE}
sim_CRnetwork <- function(J = 5, n = 105, ppA = 0.7, ppB = 0.7, psiAA = 0.3, psiBB = 0.8, pi = 0.7){
library(runjags)
# code to simulate with jags, note the use of the data block
# parameters for simulations 
# J = nb occasions
# n = nb of dyads (= N(N-1)/2 where N is the number of individuals); by default, we consider N = 15 individuals, hence N(N-1)/2 = 105 possible dyads
# ppA = detection for associated dyads
# ppB = detection for non-associated dyads
# psiAA = pr of staying associated
# psiBB = pr of staying non-associated
# pi = initial state pr
txtstring <- '					
data{
	
## States
# A = associated
# B = non-associated

## Observations (O, 1, 2 and 3 in the paper)
1 is for none of the two individuals of a dyad are seen
2 is for one individual only of the dyad is seen
3 is for the two individuals of a dyad are seen associated
4 is for the two individuals of a dyad are seen non-associated
	
# Pr(dyads in state A or B at t+1 [1 or 2 in columns] given state A or B at t [1 or 2 in rows])
px[1,1] <- psiAA # probability of staying associated
px[1,2] <- 1 - psiAA # probability of associated -> non-associated
px[2,1] <- 1 - psiBB # probability of non-associated -> associated
px[2,2] <- psiBB # probability of staying non-associated

# Pr(dyads obs in columns given dyads state in rows)
## ppA is the individual detection probability for associated dyads
## ppB is the individual detection probability for non-associated dyads
	po[1,1] <- (1-ppA) * (1-ppA)
	po[1,2] <- 2 * ppA * (1-ppA)
	po[1,3] <- ppA * ppA
	po[1,4] <- 0
	po[2,1] <- (1-ppB) * (1-ppB)
	po[2,2] <- 2 * ppB * (1-ppB)
	po[2,3] <- 0
	po[2,4] <- ppB * ppB
			
# Pr(initial states)
px0[1] <- pi # prob. of being in initial state A
px0[2] <- 1-pi # prob. of being in initial state B

# Model likelihood
	for (i in 1:n){
		
		# record states for every sampling occasion
		x1[i] <- x[i,1]
		x2[i] <- x[i,2]
		x3[i] <- x[i,3]
		x4[i] <- x[i,4]
		x5[i] <- x[i,5]
		
		# for t = 1
		x[i,1] ~ dcat(px0[1:2])
		obs[i,1] ~ dcat(po[x[i,1],1:4])

		# for t > 1
		for (t in 2:J){
			
			#-- state equation
			x[i,t] ~ dcat(px[x[i,t-1],1:2]) 
			
			#-- observation equation
			obs[i,t] ~ dcat(po[x[i,t],1:4])
							}
						}
}
model{
fake <- 0
}
'

# parameters are treated as data for the simulation step
data<-list(n=n, J=J, ppA=ppA, ppB=ppB, psiAA=psiAA, psiBB=psiBB, pi=pi)

# run jags
out <- run.jags(txtstring, data = data, monitor=c("obs","x"), sample=1, n.chains=1, summarise=FALSE)

# reformat the outputs
Simulated <- coda::as.mcmc(out)
#Simulated
#dim(Simulated)
dat <- matrix(Simulated[1:(n*J)],ncol=J)
#dat
states <- matrix(Simulated[-(1:(n*J))],ncol=J)
#states
list(dat=dat,states=states) # outputs: dat = detections/non-detections; states = underlying states
}
```

In another step, we specify the model that will be used to estimate network parameters:
```{r}
sink("sim_network_hom.txt")
cat("
model{

## States
# A = associated
# B = non-associated

## Observations (O, 1, 2 and 3 in the paper)
1 is for none of the two individuals of a dyad are seen
2 is for one individual only of the dyad is seen
3 is for the two individuals of a dyad are seen associated
4 is for the two individuals of a dyad are seen non-associated
	
# Pr(dyads in state A or B at t+1 [1 or 2 in columns] given state A or B at t [1 or 2 in rows])
px[1,1] <- psiAA 			# probability of staying associated
px[1,2] <- 1 - psiAA 		# probability of associated -> non-associated
px[2,1] <- 1 - psiBB 		# probability of non-associated -> associated
px[2,2] <- psiBB 			# probability of staying non-associated

# Pr(dyads obs in columns given dyads state in rows)
## pp is the individual detection probability
po[1,1] <- (1-pp) * (1-pp)
po[1,2] <- 2 * pp * (1-pp)
po[1,3] <- pp * pp
po[1,4] <- 0
po[2,1] <- (1-pp) * (1-pp)
po[2,2] <- 2 * pp * (1-pp)
po[2,3] <- 0
po[2,4] <- pp * pp
			
# Pr(initial states)
px0[1] <- pi 				# prob. of being in initial state A
px0[2] <- 1-pi 				# prob. of being in initial state B

# Model likelihood
for (i in 1:n){
		
	# record states for every sampling occasion
	x1[i] <- x[i,1]
	x2[i] <- x[i,2]
	x3[i] <- x[i,3]
	x4[i] <- x[i,4]
	x5[i] <- x[i,5]
		
	# for t = 1
	x[i,1] ~ dcat(px0[1:2])
	obs[i,1] ~ dcat(po[x[i,1],1:4])

	# for t > 1
	for (t in 2:J){
		#-- state equation 
		x[i,t] ~ dcat(px[x[i,t-1],1:2]) 
		#-- observation equation
		obs[i,t] ~ dcat(po[x[i,t],1:4])
			}
		}

# Priors
pp ~ dunif(0,1) # detection pr
psiAA ~ dunif(0,1) # pr of staying associated
psiBB ~ dunif(0,1) # pr of staying non-associated
pi ~ dunif(0,1) # initial state pr
}
",fill=TRUE)
sink()
```

We also consider the same model as above with heterogeneous detection probabilities:
```{r}
sink("sim_network_het.txt")
cat("
model{

## States
# A = associated
# B = non-associated

## Observations (O, 1, 2 and 3 in the paper)
1 is for none of the two individuals of a dyad are seen
2 is for one individual only of the dyad is seen
3 is for the two individuals of a dyad are seen associated
4 is for the two individuals of a dyad are seen non-associated
	
# Pr(dyads in state A or B at t+1 [1 or 2 in columns] given state A or B at t [1 or 2 in rows])
px[1,1] <- psiAA 			# probability of staying associated
px[1,2] <- 1 - psiAA 		# probability of associated -> non-associated
px[2,1] <- 1 - psiBB 		# probability of non-associated -> associated
px[2,2] <- psiBB 			# probability of staying non-associated

# Pr(dyads obs in columns given dyads state in rows)
## ppA is the individual detection probability for associated dyads
## ppB is the individual detection probability for non-associated dyads
po[1,1] <- (1-ppA) * (1-ppA)
po[1,2] <- 2 * ppA * (1-ppA)
po[1,3] <- ppA * ppA
po[1,4] <- 0
po[2,1] <- (1-ppB) * (1-ppB)
po[2,2] <- 2 * ppB * (1-ppB)
po[2,3] <- 0
po[2,4] <- ppB * ppB
			
# Pr(initial states)
px0[1] <- pi # prob. of being in initial state A
px0[2] <- 1-pi # prob. of being in initial state B

# Model likelihood
for (i in 1:n){
		
	# record states for every sampling occasion
	x1[i] <- x[i,1]
	x2[i] <- x[i,2]
	x3[i] <- x[i,3]
	x4[i] <- x[i,4]
	x5[i] <- x[i,5]
		
	# for t = 1
	x[i,1] ~ dcat(px0[1:2])
	obs[i,1] ~ dcat(po[x[i,1],1:4])

	# for t > 1
	for (t in 2:J){
		#-- state equation 
		x[i,t] ~ dcat(px[x[i,t-1],1:2]) 
		#-- observation equation
		obs[i,t] ~ dcat(po[x[i,t],1:4])
			}
		}

# Priors
ppA ~ dunif(0,1) # detection pr
ppB ~ dunif(0,1) # detection pr
psiAA ~ dunif(0,1) # pr of staying associated
psiBB ~ dunif(0,1) # pr of staying non-associated
pi ~ dunif(0,1) # initial state pr
}
",fill=TRUE)
sink()
```

## 3. Simulations: Scenarios with homogeneous detection probabilities

Now we proceed with the simulations. First, we consider the situation where detection probabilities are homogeneous irrespective of the status of the dyads. We define the scenarios we would like to investigate:

* scenarios on the detection probability $p_A = p_B$: 0.3, 0.8;

* scenarios on $\pi$: 0.2, 0.7;

* scenarios on $\psi_{AA}$: 0.1, 0.4, 0.9;

* scenarios on $\psi_{BB}$: 0.1, 0.4, 0.9.

Therefore, in total we have 36 scenarios.

```{r}
grid <- expand.grid(pp=c(0.3,0.8),pi=c(0.2,0.7),psiAA=c(0.1,0.4,0.9),psiBB=c(0.1,0.4,0.9))
grid
```

Let us run the simulations, with 100 Monte Carlo iterations. These simulations take a few hours, we do not recommend to run them. For convenience, all the results are stored in the object `simul_network_index36_sim100homogeneous.RData` that we provide. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE, eval=FALSE}
# nb of monte carlo iterations
nb_simulations <- 100

# matrix to store results with estimated values for pp, psiAA, psiBB, pi
res <- array(NA,dim=c(nrow(grid),nb_simulations,4))

# run simulation
for (index in 1:nrow(grid)){ # go through grid of scenarios
		for (i in 1:nb_simulations){
 
# 1. simulate
 
pp <- grid[index,1]
pi <- grid[index,2]
psiAA <- grid[index,3]
psiBB <- grid[index,4]

sim_data <- sim_CRnetwork(J=5,n=105, ppA = pp, ppB = pp, psiAA = psiAA, psiBB = psiBB, pi = pi)
dat <- sim_data[[1]]
states <- sim_data[[2]]

# 2. estimation

# initial values
init1 <- list(psiAA=grid[index,3],pp=grid[index,1],x=states)
inits <- list(init1)
 
# data
jags.data <- list(obs = dat, n = nrow(dat), J = ncol(dat)) 
 
# nb iterations
ni <- 2000
# nb burn-in
nb <- 1000
# nb thin
nt <- 1
# nb chains
nc <- 1

# parameters to be monitored
parameters_sim <- c("psiAA","psiBB","pi","pp","x1","x2","x3","x4","x5")

# call JAGS from R
mod <- jags(jags.data, inits, parameters_sim, 'sim_network.txt', n.chains = nc, n.thin = nt, 
n.iter = ni, n.burnin = nb, working.directory = getwd())

res[index,i,1] <- mean(mod$BUGSoutput$sims.matrix[,'pp']) # detection
res[index,i,2] <- mean(mod$BUGSoutput$sims.matrix[,'psiAA']) # associated
res[index,i,3] <- mean(mod$BUGSoutput$sims.matrix[,'psiBB']) # non-associated
res[index,i,4] <- mean(mod$BUGSoutput$sims.matrix[,'pi']) # prop of associated
   
}
 
}
save(res,file='simul_network_index36_sim100homogeneous.RData')
```

Let us post-process the results by computing relative bias (in percent) for all parameters:
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
load("simul_network_index36_sim100homogeneous.RData")
bias_param <- matrix(NA,nrow(grid),4)
for(i in 1:nrow(grid)){
	for (j in 1:4){ 
		bias_param[i,j] <- (mean(res[i,,c(1,4,2,3)[j]]) - grid[i,j])/grid[i,c(1,3,4,2)[j]]*100
	}
}
res_bias <- round(cbind(1:nrow(bias_param),grid,bias_param),2)
colnames(res_bias) <- c('scenario',names(grid),'bias_pp','bias_pi','bias_psiAA','bias_psiBB')
```

The results are given in the following table, with in the first column the scenarios labels, in columns 2-5 the simulation parameters and in columns 6-9 the relative bias:
```{r}
knitr::kable(res_bias)
```


## 4. Simulations: Scenarios with heterogeneous detection probabilities

Second, we consider the situation where detection probabilities are heterogeneous depending on the status of the dyads. We define the scenarios we would like to investigate:

* scenarios on the detection probability: $p_A = 0.3$ and $p_B = 0.8$ vs. $p_A = 0.8$ and $p_B = 0.3$

* scenarios on $\pi$: 0.2, 0.7;

* scenarios on $\psi_{AA}$: 0.1, 0.4, 0.9;

* scenarios on $\psi_{BB}$: 0.1, 0.4, 0.9.

Therefore, in total we have 36 scenarios.

```{r}
grid3 <- expand.grid(pi=c(0.2,0.7),psiAA=c(0.1,0.4,0.9),psiBB=c(0.1,0.4,0.9))
grid2 <- cbind(pA = 0.3, pB = 0.8, grid3)
grid1 <- cbind(pA = 0.8, pB = 0.3, grid3)
grid <- rbind(grid2,grid1)
grid
```

Let us run the simulations, with 100 Monte Carlo iterations. These simulations take a few hours, we do not recommend to run them. For convenience, all the results are stored in the object `simul_network_index36_sim100heterogeneous.RData` that we provide. 

```{r message=FALSE, warning=FALSE, paged.print=FALSE, eval=FALSE}
# nb of monte carlo iterations
nb_simulations <- 100

# matrix to store results with estimated values for ppA, ppB, psiAA, psiBB, pi
res <- array(NA,dim=c(nrow(grid),nb_simulations,5))

# run simulation
for (index in 1:nrow(grid)){ # go through grid of scenarios
		for (i in 1:nb_simulations){
 
# 1. simulate
 
ppA <- grid[index,1]
ppB <- grid[index,2]
pi <- grid[index,3]
psiAA <- grid[index,4]
psiBB <- grid[index,5]

sim_data <- sim_CRnetwork(J=5,n=105, ppA = ppA, ppB = ppB, psiAA = psiAA, psiBB = psiBB, pi = pi)
dat <- sim_data[[1]]
states <- sim_data[[2]]

# 2. estimation

# initial values
init1 <- list(psiAA=grid[index,3],ppA=grid[index,1],ppB=grid[index,2],x=states)
inits <- list(init1)
 
# data
jags.data <- list(obs = dat, n = nrow(dat), J = ncol(dat)) 
 
# nb iterations
ni <- 2000
# nb burn-in
nb <- 1000
# nb thin
nt <- 1
# nb chains
nc <- 1

# parameters to be monitored
parameters_sim <- c("psiAA","psiBB","pi","ppA","ppB","x1","x2","x3","x4","x5")

# call JAGS from R
mod <- jags(jags.data, inits, parameters_sim, 'sim_network_het.txt', n.chains = nc, n.thin = nt, 
n.iter = ni, n.burnin = nb, working.directory = getwd())

res[index,i,1] <- mean(mod$BUGSoutput$sims.matrix[,'ppA']) # detection
res[index,i,2] <- mean(mod$BUGSoutput$sims.matrix[,'ppB']) # detection
res[index,i,3] <- mean(mod$BUGSoutput$sims.matrix[,'psiAA']) # associated
res[index,i,4] <- mean(mod$BUGSoutput$sims.matrix[,'psiBB']) # non-associated
res[index,i,5] <- mean(mod$BUGSoutput$sims.matrix[,'pi']) # prop of associated
   
}
 
}
save(res,file='simul_network_index36_sim100heterogeneous.RData')
```

Let us post-process the results by computing relative bias (in percent) for all parameters:
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
load("simul_network_index36_sim100heterogeneous.RData")
bias_param <- matrix(NA,nrow(grid),5)
for(i in 1:nrow(grid)){
	for (j in 1:5){ 
		bias_param[i,j] <- (mean(res[i,,c(1,2,5,3,4)[j]]) - grid[i,j])/grid[i,c(1,2,5,3,4)[j]]*100
	}
}
res_bias <- round(cbind(1:nrow(bias_param),grid,bias_param),2)
colnames(res_bias) <- c('scenario',names(grid),'bias_ppA','bias_ppB','bias_pi','bias_psiAA','bias_psiBB')
```

The results are given in the following table, with in the first column the scenarios labels, in columns 2-6 the simulation parameters and in columns 7-11 the relative bias:
```{r}
knitr::kable(res_bias)
```

